\ProjectName{}: Data Imputation as a Query Optimization

### 2017-02-22 paper/algorithm.tex
paper: some typos, clarifications, fixes

Our optimizer searches a restricted space (Sec.~\ref{sec:search-space}) of query plans containing imputation operators (Sec.~\ref{sec:operators}) for a plan that minimizes a metric (Sec.~\ref{sec:cost-model}) which combines the runtime cost of the query and the estimated quality of the results.
This plan is subject to two constraints (Sec.~\ref{sec:placement}).
First, it must not emit any tuples which contain null values, regardless of the state of the base tables.
Second, the traditional relational operators (selection $\sigma$, projection $\pi$, join $\bowtie$, and group-by/aggregate $g$) must never observe a null value in any attribute that they operate on directly.
Our planning algorithm is agnostic to the type of imputation used (Sec.~\ref{sec:imputation}), as it is treated as a black-box operation.

@@

\subsection{Cost model}
\label{sec:cost-model}
The cost for a plan $q$ is expressed as a tuple of $\langle \textsc{Loss}(q), \textsc{Time}(q) \rangle$.
$\textsc{Time}(q)$ is an estimate of the runtime of a query $q$ derived from table statistics and selectivity estimation of the query predicates. $\textsc{Loss}(q)$ is an estimate of the amount of error introduced by a sequence of imputation procedure. Our
cost model averages over the losses from each such operator. We note that our loss calculation constrains $\textsc{Loss}(q)$ to $[0, 1]$, while $\textsc{Time}(q)$ can range within $(0, \infty)$.

@@

$\alpha = 1.0$ means that the optimal query should have the highest performance (at the expense of quality), while $\alpha=0.0$ means that the optimal query should lose
as little information as possible (at the expense of performance).

The computation of $\textsc{Time}(q)$ and $\textsc{Loss}(q)$ relies on having an accurate cardinality estimate of $q$ and its sub-queries.
These cardinality estimates are impacted not just by filtering or joining, as in the traditional relational calculus, but also by the imputation operators.
For example, a drop operator will reduce the cardinality of the result while an impute operator will maintain the same cardinality as the input.
For simplicity, each of the logical nodes in a query plan points to a set of histograms.
When the optimizer creates a new query plan, it copies the histograms of the sub-plans and modifies them as necessary to account for new operation in the plan.
Algorithm~\ref{algo:histogram-transformation} describes the process of generating new histograms from sub-plans.

For each sub-query, we keep track of the estimated number of
null values in a column  ($\textsc{Missing}(c, Q)$) by using the histograms associated with
the sub-query. We use this estimate to compute the set $\text{Loss}(Q)$ as follows.

\begin{align*}
  \textsc{Loss}(q) = \begin{cases}
    \{1\} \cup \textsc{Loss}(q') & q = \delta_C(q') \\
    \{\frac{1}{\sqrt{ \textsc{Attr}(q') * \textsc{Cardinality}(q') }} \} \cup \textsc{Loss}(q')  & q = \mu_C(q') \\
    \textsc{Loss}(q_1') \cup \textsc{Loss}(q_2') & q = q_1' \bowtie_\psi q_2' \\
    \textsc{Loss}(q') & q = \sigma_\phi(q'), \\ & q = \pi_C(q') \\
  \end{cases}
\end{align*}

@@

In order to avoid having the time component of our cost model dominate the information loss 
component, we scale the former by an arbitrary constant.

@@

\begin{figure}
  \begin{align*}
    \llbracket Q[T] \rrbracket &= \begin{cases}
      P & (T, P) \in Q \\
      \emptyset & \text{otherwise} \\
    \end{cases} \\
    \llbracket Q[T] \lhd p \rrbracket &= \begin{aligned} Q \cup \{(T,\ P') \} \end{aligned}\\
                                            \text{ where } \\
                                           P' &= \left\{q ~|~ \begin{aligned}
                                                                     q \in \llbracket Q[T] \rrbracket \cup \{p\}, \\
                                                                     \textsc{Dirty}(q) = \textsc{Dirty}(p),\\
                                                                     \forall q' \in \llbracket Q[T] \rrbracket.  \doms{q} {q'} 
                                                                   \end{aligned}
                                           \right\}
  \end{align*}
  \caption{Semantics of the plan cache.}
  \label{fig:semantics-plan-cache}
\end{figure}

@@

The planner operates as follows.
First, it collects the set of attributes which are used by the operators in the plan or which are visible in the output of the plan.
This set will be used to determine which attributes are imputed.
Then, it constructs a plan cache. For each table used in the query, any available selections are pushed to the leaves and
a search over imputation operations produces various possible plans, which are added to the cache. If no selections are available, a 
simple scan is added to the plan cache. Join order is jointly optimized with imputation placements and a set of
plans encompassing all necessary tables is produced. This set is now extended for any necessary grouping and aggregation
operations. After a similar step has been taken for projections, the planner returns the cheapest plan from the cache.

@@

The flexibility of this non-parametric imputation approach aligns well with the spirit of
query optimization, which aims to reconcile performance with a declarative interface to data
manipulation. By removing concerns for imputation, users executing queries in \ProjectName{}
don't need to consider the potential nature of missing data.

@@

\subsection{Complexity}
\todo{this section needs to be cleaned up/clarified}
Our optimization algorithm builds off the approach taken by System R \cite{blasgen1981system}, therefore our algorithm still operates in exponential time. Indeed, 
note that if we remove our restriction on types of imputation (i.e. minimal vs maximal), and allow any arbitrary subset of attributes to be imputed,
then every single operator in the query plan has a number of imputations exponential in the number of dirty columns. Our restriction, instead, increases the number
of plans (in the worst case) at each operator by a factor of 3. Of course, this implies that in the worst case (where all dirty sets tracked are distinct throughout the query plan),
we explore a number of plans that further scales the number of plans in the original algorithm by an exponential factor.

>>>>

Our optimizer searches a restricted space (Sec.~\ref{sec:search-space}) of query plans containing imputation operators (Sec.~\ref{sec:operators}) for an optimal plan (Sec.~\ref{sec:cost-model}) given an estimate of the runtime of the query, information lost through imputation operations, and a parameter trading off the two.

This plan is subject to two constraints (Sec.~\ref{sec:placement}).
First, it must not emit any tuples which contain null values, regardless of the state of the base tables.
Second, the traditional relational operators (selection $\sigma$, projection $\pi$, join $\bowtie$, and group-by/aggregate $g$) must never observe a null value in any attribute that they operate on directly.
Our planning algorithm is agnostic to the type of imputation used (Sec.~\ref{sec:imputation}), as it is treated as a black-box operation.

@@

\subsection{Cost model}
\label{sec:cost-model}
The cost for a plan $q$ is expressed as a tuple of $\langle \textsc{Loss}(q), \textsc{Time}(q) \rangle$.
$\textsc{Time}(q)$ is an estimate of the runtime of a query $q$ derived from table statistics and selectivity estimation of any query predicates. $\textsc{Loss}(q)$ is an estimate of the amount of information lost through the sequence of imputation procedures used. We note that our loss calculation constrains $\textsc{Loss}(q)$ to $[0, 1]$, while $\textsc{Time}(q)$ can range within $(0, \infty)$.

@@

$\alpha = 1.0$ means that the optimal query should have the fastest runtime (at the expense of quality), while $\alpha=0.0$ means that the optimal query should lose
as little information as possible (at the expense of performance).

The computation of $\textsc{Time}(q)$ and $\textsc{Loss}(q)$ relies on having an accurate cardinality estimate of a query plan $q$ and its sub-plans.
These cardinality estimates are impacted not just by filtering or joining, as in the traditional relational calculus, but also by the imputation operators.
For example, a drop operator will reduce the cardinality of the result while an impute operator will maintain the same cardinality as the input.
For simplicity, each of the logical nodes in a query plan points to a set of histograms.
When the optimizer creates a new query plan, it copies the histograms of the sub-plans and modifies them as necessary to account for new operation in the plan.
Algorithm~\ref{algo:histogram-transformation} describes the process of generating new histograms from sub-plans.

For each sub-plan, we keep track of the estimated number of
null values in a column  ($\textsc{Missing}(c, Q)$) by using the histograms associated with
that plan. We use this estimate to compute $\text{Loss}(Q)$ as follows.

\begin{align*}
  \textsc{Losses}(q) &= \begin{cases}
    \{1\} \cup \textsc{Loss}(q') & q = \delta_C(q') \\
    \{\frac{1}{\sqrt{ \textsc{Attr}(q') * \textsc{Cardinality}(q') }} \} \cup \textsc{Losses}(q')  & q = \mu_C(q') \\
    \textsc{Losses}(q_1') \cup \textsc{Losses}(q_2') & q = q_1' \bowtie_\psi q_2' \\
    \textsc{Losses}(q') & q = \sigma_\phi(q'), \\ & q = \pi_C(q') \\
  \end{cases}\\
  \textsc{Loss}(q) &= \frac{\Sigma_{l \in \textsc{Losses}(q)} l}{| \textsc{Losses}(q)|}
\end{align*}

@@

@@

\begin{figure}
  \begin{align*}
    \llbracket Q[T] \rrbracket &= \begin{cases}
      P & (T, P) \in Q \\
      \emptyset & \text{otherwise} \\
    \end{cases} \\
    \llbracket Q[T] \lhd p \rrbracket &= \begin{aligned} Q \cup \{(T,\ P') \} \end{aligned}\\
                                            \text{ where } \\
                                           P' &= \left\{q ~|~ \begin{aligned}
                                                                     q \in \llbracket Q[T] \rrbracket \cup \{p\}, \\
                                                                     \textsc{Dirty}(q) = \textsc{Dirty}(p),\\
                                                                     \forall q' \in \llbracket Q[T] \rrbracket.  \doms{q} {q'} 
                                                                   \end{aligned}
                                           \right\}
  \end{align*}
  \caption{Semantics of a plan cache $Q$}
  \label{fig:semantics-plan-cache}
\end{figure}

@@

The planner operates as follows.
First, it collects the set of attributes which are used by the operators in the plan or which are visible in the output of the plan.
This set will be used to determine which attributes are imputed.
Then, it constructs a plan cache. For each table used in the query, any available selections are pushed to the leaves and
a search over imputation operations produces various possible plans, which are added to the cache. If no selections are available, a 
simple scan is added to the plan cache. Join order is jointly optimized with imputation placements and a set of
plans encompassing all necessary tables is produced. This set is now extended for any necessary grouping and aggregation
operations. After a similar step has been taken for projections, the planner now contains the plan frontier (i.e. the
best possible set from which to pick our final plan). The final step in the planner is to find all plans that are
$\alpha$-bound and return the one that has the lowest runtime: the $\alpha$-bound optimal plan.

@@

The flexibility of this non-parametric imputation approach aligns well with the spirit of
query optimization, which aims to reconcile performance with a declarative interface to data
manipulation. By removing concerns for imputation, users executing queries in \ProjectName{}
do not need to consider the potential nature of missing data.

@@

\subsection{Complexity}
\todo{this section needs to be cleaned up/clarified}
Our optimization algorithm builds off the approach taken by System R \cite{blasgen1981system}, therefore our algorithm still operates in exponential time. Indeed, 
note that if we remove our restriction on imputation operators (i.e. minimal vs maximal), and allow any arbitrary subset of attributes to be imputed,
then every single operator in the query plan has a number of imputations exponential in the number of dirty columns. Our restriction, instead, increases the number
of plans (in the worst case) at each operator by a factor of 3. Of course, this implies that in the worst case (where all dirty sets tracked are distinct throughout the query plan),
we explore a number of plans that further scales the number of plans in the original algorithm by an exponential factor.

### 2017-02-22 paper/conclusion.tex
paper: some typos, clarifications, fixes

\section{Conclusion}
As shown in \ProjectName{}, missing values and their imputation can be successfully integrated into the relational calculus and
existing cost-based plan optimization frameworks. We implement imputation actions, such as dropping or imputing values with a machine
learning technique, as operators in the algebra and use a simple, but effective, cost model to consider tradeoffs in 
information loss and runtime. We showed that different values for the tradeoff parameter can yield substantially
different plans, allowing the user to express their preferences for performance on a per-query basis.
Our experiments with \textit{CDC} and \textit{freeCodeCamp} survey-based datasets
showed that \ProjectName{} can be used successfully with real-world data, expressing standard SQL queries over data with missing values. 
We craft a series of realistic queries, meant to resemble the kind of questions an analyst might ask during
the data exploration phase. The plans selected for each query execute in a fraction of the time compared to the standard approach of imputing on the entire base tables and
then formulating queries. Furthermore, the difference in query results between the two approaches is
shown to be minimal. Data analysts need not commit to a specific imputation strategy and can instead
vary this across queries.

We highlight the long history of dealing with missing data both in the statistical learning and database communities.
Similarly to existing work, we consider the impact of null values in databases and develop a simple set of invariants to 
successfully plan around them. In contrast to the statistical learning work, our emphasis is not on the specific algorithm
used to impute but rather on the timing of imputation in the execution of a query. In contrast to existing database work,
we incorporate imputation into a cost-based optimizer and hide any details
regarding missing values inside the system, allowing users to use traditional SQL and engage in normal workloads.
While prior work has incorporated operations such as prediction into their databases, this has been domain-specific
in some cases, and in others, they have not integrated these operations into the planner. The novelty
of our contribution lies in the formalization of missing value imputation for query planning, which results in performance
gains over traditional approaches. This approach acknowledges that different users performing different queries on the same
dataset will likely have varying imputation needs and that execution plans should appropriately reflect that variation.

@@

\textbf{Imputation operators:} We explored a subset of possible imputation operators.
In particular, we explored the two variants of the drop and impute operations. The system
could be extended to consider a broader family of operators. As the search space grows,
there will likely be additional steps needed in order to avoid sub-optimal plans.

>>>>

\section{Conclusion}
As shown in \ProjectName{}, missing values and their imputation can be successfully integrated into the relational calculus and
existing cost-based plan optimization frameworks. We implement imputation actions, such as dropping or imputing values with a machine
learning technique, as operators in the algebra and use a simple, but effective, cost model to consider tradeoffs in 
information loss and runtime. We showed that different values for the tradeoff parameter can yield substantially
different plans, allowing the user to express their preferences for performance on a per-query basis.
Our experiments with \textit{CDC} and \textit{freeCodeCamp} survey-based datasets
showed that \ProjectName{} can be used successfully with real-world data, expressing standard SQL queries over data with missing values. 
We craft a series of realistic queries, meant to resemble the kind of questions an analyst might ask during
the data exploration phase. The plans selected for each query execute in a fraction of the time compared to the standard approach of imputing on the entire base tables and
then formulating queries. Furthermore, the difference in query results between the two approaches is
shown to be small in all queries considered. Data analysts need not commit to a specific imputation strategy and can instead
vary this across queries.

We highlight the long history of dealing with missing data both in the statistical learning and database communities.
Similarly to existing work, we consider the impact of null values in databases and develop a simple set of invariants to 
successfully plan around them. In contrast to existing statistical learning work, our emphasis is not on the specific algorithm
used to impute but rather on the placement of imputation steps in the execution of a query. In contrast to existing database work,
we incorporate imputation into a cost-based optimizer and hide any details
regarding missing values inside the system, allowing users to use traditional SQL and engage in normal workloads.
While prior work has incorporated operations such as prediction into their databases, this has been domain-specific
in some cases, and in others, they have not integrated these operations into the planner. The novelty
of our contribution lies in the formalization of missing value imputation for query planning, which results in performance
gains over traditional approaches. This approach acknowledges that different users performing different queries on the same
dataset will likely have varying imputation needs and that execution plans should appropriately reflect that variation.

@@

\textbf{Imputation operators:} We explored a subset of possible imputation operators.
In particular, we explored two variants of the \textit{Drop} and \textit{Impute} operations. The system
could be extended to consider a broader family of operators. As the search space grows,
there will likely be additional steps needed in order to avoid sub-optimal plans.

### 2017-02-22 paper/experiments.tex
paper: some typos, clarifications, fixes

\subsection{Data sets} \label{subsec:datasets}
\subsubsection{Center for Disease Control: Surveys}
For our first set of experiments, we use survey data collected by the 
Center for Disease Control (CDC) in the United States. In particular, we
experiment on a set of tables collected as part of the National
Health and Nutrition Examination Survey (NHANES), a series of studies
conducted by the CDC on a national sample of several thousand individuals\cite{cdc-data}.
The data consists on survey responses, physical examinations, and laboratory
results, amongst others. Our dataset corresponds to the 2013-2014 NHANES
and is available for direct download on the CDC's organization site on 
kaggle.com, a popular data science website.

@@

All tables used consists of 10 attributes. The laboratory and examinations data,
referred to as \textit{labs} and \textit{exams} in our queries, consists of 9813 rows,
while the demographics data (\textit{demo}) has 10175 rows.

@@

Queries 1-3 are on the CDC data. Query~\ref{q1} calculates the average weight for users based on their income data.
To do so, it joins demographics and physical examination data. Query~\ref{q2}
compares cholesterol levels for individuals who are both ends of the income
spectrum and who satisfy a minimum weight requirement. Query~\ref{q3}
extracts the maximum blood lead levels for subjects who satisfy a particular
waist circumference requirement.

Queries 4-7 are on the FCC data. In Query~\ref{q4}, we calculate the average income for survey participants,
based on their bootcamp attendance. Query~\ref{q5} estimates the average
age of women from the Uniter States who participated. Query~\ref{q6} 
calculates the average amount of money survey participants with student debt destined to
learning based on their school degree. Finally, in Query~\ref{q7}, we join the FCC data with a reference table provided
by the World Bank which summarizes GDP per Capita across various
countries\cite{worldbank-data}. The query calculates the average GDP-per-capita for participants based on whether or not they
attended a bootcamp.

@@

%\todobox{actually run experiments and rewrite this section}{
All experiments were run on a Macbook Pro with 2 2.9 GHz Intel core i5 processors, 16 GB of main memory,  running OS X El Capitan.

Figure~\ref{fig:runtimes} provides a summary of the performance results. Each query is run in two configurations: quality-optimized and runtime-optimized.
To provide comparison, we also include a table with the average running time required to perform imputation at each of the base tables used in our analysis.
Note that multiple queries perform joins between two or more base tables. In all cases, running a query through \ProjectName{} is cheaper than 
imputing a single base table, with the added flexibility of imputation tailored to the query.

@@

Table~\ref{table:smape} shows the Symmetric-Mean-Absolute-Percentage-Error (SMAPE) for \ProjectName{}'s query results when compared to 
running imputation on the base tables and executing the query on the
cleaned data. Each query was run 10 iterations in both settings and then results from the two approaches were paired up and compared tuple-wise. We average 
absolute percentage deviations across each iteration of a query, and we report this value averaged over all iterations.
We can see that optimizing for quality indeed reduces the SMAPE of query results. In general, the SMAPE relative to the
base-imputation approach are low in all cases.

@@

In many cases, applying the imputation step at the base table simply is not an option. To illustrate the increasing difficulty of such an approach
as datasets scale, we used a simple aggregate query over the ACS dataset (Figure~\ref{query-acs}). Applying the imputation operation to the base table
is extremely expensive, in our case the process did not terminate within 20 minutes. In contrast, \ProjectName{} executes a quality-optimized version
in 9 seconds and a runtime-optimized version in 1 second. This highlights the potential benefit of using our system for early data exploration.

>>>>

\subsection{Data sets} \label{subsec:datasets}
\subsubsection{Center for Disease Control: Surveys}
For our first set of experiments, we use survey data collected by the 
Center for Disease Control and Prevention (CDC) in the United States. In particular, we
experiment on a set of tables collected as part of the National
Health and Nutrition Examination Survey (NHANES), a series of studies
conducted by the CDC on a national sample of several thousand individuals\cite{cdc-data}.
The data consists on survey responses, physical examinations, and laboratory
results, amongst others. Our dataset corresponds to the 2013-2014 NHANES
and is available for direct download on the CDC's organization site on 
kaggle.com, a popular data science website.

@@

All tables used consists of 10 attributes. The laboratory and examinations data,
referred to as \textit{labs} and \textit{exams} in our queries, consist of 9813 rows,
while the demographics data (\textit{demo}) have 10175 rows.

@@

Queries 1-3 are on the CDC data. Query~\ref{q1} calculates the average weight for users based on their income data.
To do so, it joins demographics and physical examination data. Query~\ref{q2}
compares cholesterol levels for individuals who are at both ends of the income
spectrum and who satisfy a minimum weight requirement. Query~\ref{q3}
extracts the maximum blood lead levels for subjects who satisfy a particular
waist circumference requirement.

Queries 4-7 are on the FCC data. In Query~\ref{q4}, we calculate the average income for survey participants,
based on their bootcamp attendance. Query~\ref{q5} estimates the average
age of women from the United States who participated. Query~\ref{q6} 
calculates the average amount of money survey participants with student debt destined to
learning based on their school degree. Finally, in Query~\ref{q7}, we join the FCC data with a reference table provided
by the World Bank which summarizes GDP per Capita across various
countries\cite{worldbank-data}. The query calculates the average GDP-per-capita for participants based on whether or not they
attended a bootcamp.

@@

%\todobox{actually run experiments and rewrite this section}{
All experiments were run on a Macbook Pro with two 2.9 GHz Intel core i5 processors, 16 GB of main memory,  running OS X El Capitan.

Figure~\ref{fig:runtimes} provides a summary of the performance results. The first plot shows average running times for \ProjectName{}
plans in two configurations: quality-optimized and runtime-optimized. The second plot shows, for comparison, the 
average running times required for each query when performing imputation at the base tables.
In all cases, running a query through \ProjectName{} is cheaper than 
imputing a single base table, with the added advantage of imputation tailored to the query. This latter point adds
significant flexibility to the user, who no longer needs to decide apriori what imputation strategy they have to stick to
for all their queries.

@@

Table~\ref{table:smape} shows the Symmetric-Mean-Absolute-Percentage-Error (SMAPE) for \ProjectName{}'s query results when compared to 
running imputation on the base tables and executing the query on the
cleaned data. Each query was run 10 iterations in both settings and then results from the two approaches were paired up and compared tuple-wise. We average tuple-wise
absolute percentage deviations within each iteration of a query, and we report this value averaged over all iterations.
We can see that optimizing for quality indeed reduces the SMAPE of query results. In general, the SMAPE relative to the
base-imputation approach are low in all cases, indicating that on-the-fly imputation produces similar results to
imputation at the base tables.

@@

In many cases, applying the imputation step at the base table simply is not an option. To illustrate the increasing difficulty of such an approach
as datasets scale, we used a simple aggregate query over the ACS dataset in our final experiment(Figure~\ref{query-acs}). Applying the imputation operation to the base table
is extremely expensive, in our case the process did not terminate within 20 minutes. In contrast, \ProjectName{} executes a quality-optimized version
in 9 seconds and a runtime-optimized version in 1 second. This highlights the potential benefit of using our system for early data exploration.

### 2017-02-22 paper/relatedwork.tex
paper: some typos, clarifications, fixes

In \cite{akande2015empirical}, Akande et al analyze the performance of various
multiple imputation techniques on the American Community Survey dataset. 
The computational difficulties of imputing on large base
tables are well known and can limit approaches. For example, Akande finds that
one approach (MI-GLM) is prohibitively expensive when attempting to impute on data that
includes variables with potentially large domains (ten categories in their case).
In contrast, \ProjectName{} allows users to specify a tradeoff between
information loss and computational complexity (in terms of time). Furthermore,
the query planner's imputation is guided by the requirements of each specific
query's operators, rather than requiring broad assumptions about query
workloads.

@@

BayesDB provides a great alternative for bridging the gap between
traditional databases and sophisticated modeling software. \ProjectName{}, in
contrast, aims to remain squarely in the database realm, while allowing
users to perform queries on a potentially larger subset of their data.

@@

\subsection{Forecasting and Databases}
\cite{parisi2011embedding} introduces the idea of incorporating time-series forecast operators into
databases, along with the necessary relational algebra extensions. Their work explores the theoretical
properties of forecast operators and generalizes them into a family of operators, distinguished by
the type of predictions returned. They highlight the use of forecasting for replacing missing values.
In their follow on work\cite{parisi2013temporal}, they identify various equivalence and containment
relationships when using forecast operators, which could be used to perform query plan transformations that guarantee the same result. They
explore forecast-first and forecast-last plans, which perform forecasting operations before and after traditional
relational operators, respectively.

@@

\ProjectName{} provides a declarative language, as
an extension to standard SQL, but in contrast to Fa, \ProjectName{} is not searching the space of possible
imputation models, but rather the space of query plans that incorporate imputation operators. Another major point
of distinction between \ProjectName{} and Fa is that the latter doesn't consider tradeoffs between accuracy and computation time, 
but rather returns the most accurate forecast (with some stopping criterion).

>>>>

In \cite{akande2015empirical}, Akande et al analyze the performance of various
multiple imputation techniques on the American Community Survey dataset. 
The computational difficulties of imputing on large base
tables are well known and can limit approaches. For example, Akande finds that
one approach (MI-GLM) is prohibitively expensive when attempting to impute on data that
includes variables with potentially large domains (ten categories in their case).
In contrast, \ProjectName{} allows users to specify a tradeoff between
information loss and runtime performance, allowing users to perform queries
directly on the entire dataset. Furthermore,
the query planner's imputation is guided by the requirements of each specific
query's operators, rather than requiring broad assumptions about query
workloads.

@@

BayesDB provides a great alternative for bridging the gap between
traditional databases and sophisticated modeling software. \ProjectName{}, in
contrast, aims to remain squarely in the database realm, while allowing
users to directly express queries on a potentially larger subset of their data.

@@

\subsection{Forecasting and Databases}
Parisi et al\cite{parisi2011embedding} introduce the idea of incorporating time-series forecast operators into
databases, along with the necessary relational algebra extensions. Their work explores the theoretical
properties of forecast operators and generalizes them into a family of operators, distinguished by
the type of predictions returned. They highlight the use of forecasting for replacing missing values.
In their follow on work\cite{parisi2013temporal}, they identify various equivalence and containment
relationships when using forecast operators, which could be used to perform query plan transformations that guarantee the same result. They
explore forecast-first and forecast-last plans, which perform forecasting operations before and after traditional
relational operators, respectively.

@@

Similarly to Fa, \ProjectName{} provides a declarative language, as
a subset of standard SQL. \ProjectName{}, however, is not searching the space of possible
imputation models, but rather the space of query plans that incorporate imputation operators. Another major point
of distinction between \ProjectName{} and Fa is that the latter doesn't consider tradeoffs between accuracy and computation time, but rather returns the most accurate forecast (with some stopping criterion).
